---
title: "Tooth Growth Analysis"
author: "Yadder Aceituno"
date: "January 24, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2);
```

## Loading Data and Summary

For this analysis I will use the "Tooth Growth" data. Let's load it and view a summary of the data.


```{r loading1}
data("ToothGrowth");
```

Now, let's look a summary of the data.
```{r loading2}
summary(ToothGrowth);
```

Now, let's look around the firsts rows of the data.
```{r loading3}
head(ToothGrowth);
```

You can find more information about the data clicking on this [link](https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/ToothGrowth.html)

## Comparing By Delivery Method

I will create a plot about length by delivery method to easily explore the data.

```{r delivery1}
ToothGrowth$dose <- as.factor(ToothGrowth$dose);

plot <- ggplot(aes(x=supp, y=len), data=ToothGrowth);

plot <- plot + geom_boxplot(aes(fill=supp)) + 
  xlab("Delivery Method") + 
  ylab("Tooth Length") + 
  ggtitle("Tooth Length by Delivery Method");

plot
```

Now, I will use t.test to compare the data by supplement (delivery method).

```{r delivery2}
t.test(len~supp, data = ToothGrowth);
```