---
title: "Exponential Distribution and CLT"
author: "Yadder Aceituno"
date: "January 23, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This project will show a comparission between the Exponential Distribution and CLT (Central Limit Theorem). I will calculate and compare the mean, the variance, and I will show how the distribution looks approximately normal.

## Simulations
First, I need to create 1000 simulations of the mean from 40 exponential values using lambda = 0.2.
I will use the rexp function to get those values. Over that function I will use the replicate function to get the 1000 simulations.

```{r simulations1}
# Seed for reproducibility
SEED <- 1505;
set.seed(SEED);

# Lambda value
lambda <- 0.2;
# Number of values  
n <- 40;
# Number of simulations
sim <- 1000;

# Creating the simulated values
simulated_values <- replicate(sim, mean(rexp(n, lambda)));

# Looking the firsts values in the vector
head(simulated_values);
```

## Sample Mean vs Theorical Mean
We know that we can calculate the theorical mean with this equation: $E(x) = 1/\lambda = \beta$

```{r sample_mean1}
# Calculating theorical mean.
theorical_mean <- 1 / lambda;

# Looking the value
theorical_mean;
```

Now, I will calculate the mean from my 1000 simulations.

```{r sample_mean2}
# Calculating simulations' mean
simulation_mean <- mean(simulated_values);

# Looking the value
simulation_mean;
```

We can check that both are approximately similar, the theorical mean is equal 5, and the simulations' mean is equal 5.01.

## Sample Variance vs Theorical Variance
We know that we can calculate the theorical standard deviation with this equation: $\sigma_x = \sigma / \sqrt n$
We also know that $\sigma = \beta$ for exponential distribution. So, we can calculate standard deviation with $\sigma_x = (1/\lambda) / \sqrt n$

```{r sample_variance1}
# Calculating theorical standard deviation.
theorical_standard_deviation <- (1 / lambda) / sqrt(n);

# Looking the value
theorical_standard_deviation;
```

Now, I will calculate the standard deviation from my 1000 simulations.

```{r sample_variance2}
# Calculating simulations' standard deviation
simulation_standard_deviation <- sd(simulated_values);

# Looking the value
simulation_standard_deviation;
```

With those values, we can calculate both variances.

```{r sample_variance3}
# Calculating  theorical variance
theorical_variance <- theorical_standard_deviation ^ 2;
theorical_variance;

# Calculating  simulations' variance
simulation_variance <- simulation_standard_deviation ^ 2;
simulation_variance;
```
